% run this through LaTeX with the appropriate wrapper

\chapter	{The ISODE Little Services}\label{imisc}
The {\em ISODE Little Services\/} are examples of a few simple
applications written using {\em The ISO Development Environment}.
The programs described herein should work on all systems on which the
software runs.
All of these programs have \unix/ manual entries.

The little services are based on the protocols of the same name found in
the \dod/ TCP/IP protocol suite.
There are several services:
\begin{describe}
\item[utcTime:] the universal time

\item[genTime:] the generalized time

\item[timeOfDay:]	the current date and time since the \unix/ epoch

\item[users:]	the users logged in on the system

\item[charGen:]	a character generation pattern

\item[qotd:]	a quote of the data

\item[finger:]	``fingers'' users logged in

\item[pwdGen:]	six pseudo-randomly generated (allegedly mnemonic) passwords

\item[tellUser:] sends a message to a remote user

\item[ping:] ping test for performance measurement

\item[sink:] sink test for performance measurement

\item[echo:] echo test for performance measurement
\end{describe}
Only the {\em finger\/} service takes any arguments,
the individuals on whom to report.

\section	{Implementation}
If you have access to the source tree for this release,
the directory \file{imisc/} contains the code for the responder and initiator.

\subsection	{The Initiator}
The \man imisc(1c) program is the initiator which requests the little
services.
If invoked with arguments,
it executes that exact operation and terminates.
Otherwise, it enters interactive mode,
prompting for each operation and argument until end-of-file is found.

In addition, the pseudo-operations \verb"help" and \verb"quit" do the
obvious things.

\subsection	{The Responder}
The \man imiscd(8c) program is the responder which implements the little
services.

As shown in Figure~\ref{imisc:ros} on the following page,
the ROS-based definition is very simple.
\newpage
\tagrindfile{grindimisc-1}{ROS definition of ISODE Little Services}{imisc:ros}
