% -*- LaTeX -*-		(really SLiTeX)

\documentstyle[blackandwhite,landscape,oval,pagenumbers,small,tgrind]{NRslides}

\def\emph#1{\underline{#1}}

\font\xx=cmbx10
\font\yy=cmbx7

\raggedright

\input trademark
\let\tradeNAMfont=\relax
\let\tradeORGfont=\relax

%\includeonly	{directory}


\begin{document}

\title	{ISODE INTERNALS}
\author	{Marshall T.~Rose\\ Performance Systems International, Inc.}
\date	{April 24--25, 1991}
\maketitlepage


\begin{bwslide}\thispagestyle{empty}
\part*	{AGENDA}\bf

\begin{description}
\item[TOPIC I:]		HISTORY AND INTRODUCTION

\item[TOPIC II:]	LOWER LAYER INFRASTRUCTURE

\item[TOPIC III:]	UPPER LAYER INFRASTRUCTURE

\item[TOPIC IV:]	IMPLEMENTING NEW SERVICES

\item[TOPIC V:]		NETWORK MANAGEMENT (SNMP)

\item[TOPIC VI:]	DIRECTORY SERVICES

\item[TOPIC VII:]	OTHER SERVICES
\end{description}
\end{bwslide}


\setcounter{slide}{0}
\begin{bwslide}
\ctitle	{SANITY CHECK:\\ HAVE YOU HEARD OF$\ldots$}

\begin{nrtc}
\item	CO-MODE TRANSPORT SERVICE?

\item	TOKENS, DIALOGUES, ACTIVITIES, EXCEPTIONS?

\item	ABSTRACT SYNTAX and TRANSFER NOTATION?

\item	CONTEXT MANAGEMENT, SYNTAX MATCHING?

\item	APPLICATION CONTEXTS?
\end{nrtc}
\end{bwslide}


\begin{bwslide}\small\em
\ctitle	{PURIST CHECK:\\ A CAUTIONARY TALE ABOUT OSI}

Once upon a time, in a kingdom not far from here, a king summoned two
of his advisors for a test.  He showed them both a shiny metal box
with two slots in the top, a control knob, and a lever.  ``What do you
think this is?''

One advisor, an engineer, answered first. ``It is a toaster,'' he said.
The king asked, ``How would you design an embedded computer for it?''
The engineer replied, ``Using a four-bit microcontroller, I would write
a simple program that reads the darkness knob and quantizes its
position to one of 16 shades of darkness, from snow white to coal
black. The program would use that darkness level as the index to a
16-element table of initial timer values.  Then it would turn on the
heating elements and start the timer with the initial value selected
from the table.  At the end of the time delay, it would turn off the
heat and pop up the toast.  Come back next week, and I'll show you a
working prototype.''

The second advisor, a computer scientist, immediately recognized the danger of
such short-sighted thinking.  He said, ``Toasters don't just turn bread into
toast, they are also used to warm frozen waffles.  What you see before you is
really a breakfast food cooker.  As the subjects of your kingdom become more
sophisticated, they will demand more capabilities.  They will need a breakfast
food cooker that can also cook sausage, fry bacon, and make scrambled eggs.  A
toaster that only makes toast will soon be obsolete.  If we don't look to the
future, we will have to completely redesign the toaster in just a few years.''
\end{bwslide}


\begin{bwslide}\small\em
\ctitle	{PURIST CHECK (cont.)}

``With this in mind, we can formulate a more intelligent solution to
the problem.  First, create a class of breakfast foods. Specialize
this class into subclasses: grains, pork, and poultry.  The
specialization process should be repeated with grains divided into
toast, muffins, pancakes, and waffles; pork divided into sausage,
links, and bacon; and poultry divided into scrambled eggs, hard-boiled
eggs, poached eggs, fried eggs, and various omelet classes.''

``The ham and cheese omelet class is worth special attention because it
must inherit characteristics from the pork, dairy, and poultry
classes.  Thus, we see that the problem cannot be properly solved
without multiple inheritance.  At run time, the program must create
the proper object and send a message to the object that says, `Cook
yourself.'  The semantics of this message depend, of course, on the
kind of object, so they have a different meaning to a piece of toast
than to scrambled eggs.''

``Reviewing the process so far, we see that the analysis phase has
revealed that the primary requirement is to cook any kind of breakfast
food.  In the design phase, we have discovered some derived
requirements.  Specifically, we need an object-oriented language with
multiple inheritance.  Of course, users don't want the eggs to get
cold while the bacon is frying, so concurrent processing is required,
too.''

``We must not forget the user interface.  The lever that lowers the
food lacks versatility, and the darkness knob is confusing.  Users
won't buy the product unless it has a user-friendly, graphical
interface.  When the breakfast cooker is plugged in, users should see
a cowboy boot on the screen.  Users click on it, and the message
`Booting UNIX v.~8.3' appears on the screen. (UNIX v.~8.3 should be out
by the time the product gets to the market.)  Users can pull down a
menu and click on the foods they want to cook.''
\end{bwslide}


\begin{bwslide}\small\em
\ctitle	{PURIST CHECK (cont.)}

``Having made the wise decision of specifying the software first in the design
phase, all that remains is to pick an adequate hardware platform for the
implementation phase.  An Intel 80386 with 8MB of memory, a 30MB hard disk, and
a VGA monitor should be sufficient.  If you select a multitasking, object
oriented language that supports multiple inheritance and has a built-in GUI,
writing the program will be a snap.  (Imagine the difficulty we would have had
if we had foolishly allowed a hardware-first design strategy to lock us into a
four-bit microcontroller!).''

The king had the computer scientist thrown in the moat, and they all lived
happily ever after.

\raggedright
---unattributed, reported by Manny Farber
\end{bwslide}


\begin{bwslide}
\ctitle	{AN IMPORTANT CONVENTION}

\begin{nrtc}
\item	A LOT OF PROCEDURE NAMES ARE PRESENTED IN THIS TALK

\item	TO FIND OUT MORE:
    \begin{nrtc}
    \item	SIMPLY LOOK IN THE \emph{INDEX}

    \item	OF THE APPROPRIATE VOLUME OF THE MANUAL
    \end{nrtc}
\end{nrtc}
\end{bwslide}


\begin{bwslide}
\ctitle	{DON'T BE BASHFUL}

\begin{nrtc}
\item	YOU AREN'T EXPECTED TO BE AN EXPERT ON THINGS LIKE
    \begin{nrtc}
    \item	OSI ADDRESSING, or

    \item	THE ISODE
    \end{nrtc}
    SO ASK QUESTIONS$\ldots$

\item	ALSO, THIS IS AN INTERNALS COURSE
    \begin{nrtc}
    \item	SO WE'RE NOT GOING TO DO COMPARISONS WITH OTHER SERVICES
		(e.g., DIRECTORY v. DNS)
    \end{nrtc}
\end{nrtc}
\end{bwslide}


\include	{history}
\include	{lower}
\include	{upper}
\include	{building}
\include	{snmp}
\include	{directory}
\include	{others}

\end{document}
