% run this through LaTeX with the appropriate wrapper

\input lcustom

\documentstyle[12pt]{article}

\begin{document}

\section*	{Keywords}
Open systems interconnection,
protocol migration,
transport service,
interface translation,
magic boxes.


\section*	{Biographies}
{\bf Marshall T.~Rose} received the B.S., M.S., and Ph.D. degrees in
Information and Computer Science from the University of California, Irvine,
in~1981, 1983, and~1984 respectively.
He has worked on formal methods of specification
and verification of computer network protocols,
models of parallel and distributed computation,
and computer-based message systems.
His current interests are in high-performance distributed memory systems in
support of abstractions for interprocess communication,
development of an environment for the development of advanced application
protocols,
and visualization techniques for software design, implementation, and testing.

He currently works in the Computer Science Laboratory at
Northrop Research and Technology Center,
and is also an Adjunct Assistant Professor at the University of Delaware.

\vspace{0.25in}

{\bf Dwight E.~Cass} has worked for the previous 14~years in systems-level
computer science for several commercial and research organizations,
including Digital Equipment Corporation,
Zilog,
and the USC/Information Sciences Institute.
For the last 8~years,
he has focused his activities on distributed processing systems
working specifically with local area networks,
process control and management systems.
His current interests are in factory floor networking concepts
and robotic technology research.

He currently works in the Automation Sciences Laboratory at Northrop Research
and Technology Center.
\end{document}
