% run this through LaTeX

\input lcustom

\documentstyle[11pt,sfwmac,titlepage]{article}

\advance\textwidth by0.75in
\advance\oddsidemargin by-0.375in
\advance\evensidemargin by-0.375in

\pagestyle{reprint}
\def\reprintext{{\scriptsize\sf
	Reprinted from
	{\em Computer Networks and ISDN Systems}, 12(3) {\oldstyle 1986}}}
\let\reprintstring=\empty

\begin{document}

\title{OSI Transport Services\\ on top of the TCP}
\author{Marshall T.~Rose\and Dwight E.~Cass}
\date{\vskip-1.25em
	Northrop Research and Technology Center\\
	One Research Park\\
	Palos Verdes Peninsula, CA  90274\\
	USA\vskip1.5em
	\ifdraft \versiondate/\\ \tt Draft Version \versiontag/\else \today\fi}
\maketitle

\begin{abstract}
The ARPA Internet community has a well-developed,
mature set of transport and internetwork protocols,
which are quite successful in offering network and transport services to
end-users.
The protocols provide ``open system interconnection'' (OSI) for users,
and a robust competitive market is supplied by a growing number of vendors.
The CCITT and ISO have defined various network, transport, session, presentation,
and application recommendations and standards which have been adopted by the
international community and numerous vendors.
To the largest extent possible,
it is desirable to offer the higher level services (i.e., session and above)
in networks based on the ARPA protocols,
without disrupting existing facilities.
It is also desirable to develop expertise with ISO and CCITT protocols while
utilizing a robust set of transport services.
Finally,
given the probable, long-term dominance of the international standards,
it is desirable to design a graceful transition strategy from networks based
on the ARPA protocols to OSI-based networks.

In this paper,
we suggest an exercise in interoperability which achieves all three of these
goals.
Our solution is practical in the sense that it:
can be easily implemented in a cost-effective manner in the short-term;
permits application-designers to work with an OSI applications-framework,
while utilizing existing robust transport and network protocols;
and,
provides the basis for a far-reaching migration strategy to the OSI protocols.
\end{abstract}

\newpage\thispagestyle{empty}
\tableofcontents
\footnotetext[0]{\hskip -\parindent
This document (version \versiontag/)
was \LaTeX set \today\ with \fmtname\ v\fmtversion.
This document is also referenced as NRTC Technical Paper \#700.
This work was supported in part by Northrop's Independent Research and
Development program.}

\makeatletter
\let\s@ction=\section
\def\section{\let\section=\s@ction \banner\section}
\makeatother

\include{section1}
\include{section2}
\include{section3}
\include{section4}
\include{section5}
\include{section6}

\appendix
\include{appendix1}
\include{appendix2}

\bibliography{bcustom,networking}
\bibliographystyle{alpha}

\showsummary

\end{document}
