% run this through LaTeX with the appropriate wrapper

\chapter	{Introduction}
This document is {\em The User's Handbook\/} for the
PSI White Pages Pilot.
The goal of \thebook/ is to provide a user with enough information to be able
to make effective use of the white pages service.
In practical terms,
this means that \thebook/ provides information on how
to use the white pages from your host.

The OSI Directory is used to provide the white pages service.
\thebook/ is not intended as a tutorial nor a detailed description
of the OSI Directory.
However,
in order to make effective use of the white pages,
it is necessary to understand some rudimentary Directory concepts.

Your comments are welcome!
The OSI Directory is a new, complex technology.
Although \thebook/ attempts to be straight-forward it probably doesn't
succeed all the time.
If you have comments on this document, send them to the Internet
mailbox
\begin{quote}\begin{verbatim}
wpp-camayocs@uu.psi.com
\end{verbatim}\end{quote}
so that \thebook/ can be improved.

If you are interested in sharing about experiences with other users of the
white pages, drop a note to
\begin{quote}\begin{verbatim}
wpp-users-request@psi.com
\end{verbatim}\end{quote}
and ask to be added to the
\begin{quote}\begin{verbatim}
wpp-users@psi.com
\end{verbatim}\end{quote}
discussion group.

\newpage

\section*	{Related Documentation}
In theory,
\thebook/ is the only reference needed by all but the most sophisticated of
users.
However,
if you're interested in finding out more about the PSI White Pages
Pilot Project and the underlying mechanisms and technology,
there are a few documents you might want to look at.

The white paper
{\em An Introduction to a NYSERNet White Pages Pilot Project\/}
introduces the goals and phases of the pilot project.
In addition,
site administrators are provided with a document called
{\em PSI White Pages Pilot Project: Administrator's Guide},
or simply \theguide/.

The research note {\em The Design of QUIPU\/} describes the design of the
software used for the pilot project,
whilst {\em Volume Five\/} of the ISODE User's Manual,
henceforth termed \volfive/,
describes the implementation of the software.

Your site administrator has access to all of these documents and can provide
them to you as requested.

\newpage

\input{wp-intro}

\input{dir-intro}

\section	{Where to turn for help}
The software supporting the pilot project is the QUIPU implementation
of the OSI Directory,
which is a part of the ISO Development Environment (ISODE).

QUIPU was originally developed as a part of the INCA project
(under the auspices of the ESPRIT initiative of the EEC).
The Inca of Peru did not have writing.
Instead,
they stored information on strings,
carefully knotted in a specific manner with colored thread,
and attached to a larger rope.
Such a device was known as a {\em Quipu\/}
(pronounced {\em kwip-ooo}).
The encoding was obscure,
and could only be read by selected trained people:
the {\em Quipucamayocs}.
The Quipu was a key component of Inca society,
as it contained information about property and locations throughout
the extensive Inca empire.

The pilot sponsors and the administrators of the participating organizations
form the {\em camayocs\/} for the pilot project.
There is a special mailing list
\begin{quote}\small\begin{verbatim}
wpp-camayocs@uu.psi.com
\end{verbatim}\end{quote}
which each \camayoc/ belongs to.

If you need help with the white pages service,
you should first contact your site administrator.
If you are unable to do so,
then you should contact another \camayoc/ or drop a note to the
\verb"wpp-camayocs" list.
