% run this through LaTeX with the appropriate wrapper

\section	{The White Pages Service}
A natural function of computer networks is to form the {\em infrastructure\/}
between the users they interconnect.
For example,
the electronic mail service offered by computer networks provides a means for
users to collaborate towards some common goal.
In the simplest cases,
this collaboration may be solely for the dissemination of information.
In other cases,
two users may work on joint research project,
using electronic mail as their primary means of communication.

Most network services are based on the implicit assumption that each user can
supply {\em infrastructural information} to 
facilitate information transfers through the network.
For example,
electronic mail services expect that an originator can supply 
addressing information 
for all the intended recipients.
It is not necessarily the task of electronic mail, per se,
to provide this infrastructural information to the user.

This model works fine in small environments,
particularly those where infrastructural information is not difficult to 
obtain and remember.
However,
the model does not scale well.
Consider the case when the membership of a network consists of hundreds of
thousands of users belonging to thousands of organizations.
It is no longer reasonable for a single user to provide this information,
except in very limited circumstances.
Further,
it is likely that some of the information changes frequently,
due to personnel and other resource movement.
The goal of a {\em white pages\/} service is to 
provide the necessary information, and to mask the complexity of the
infrastructural information.

From the user's perspective,
the PSI White Pages Pilot Project focuses solely on providing
infrastructural information dealing with human users.%
\footnote{The white pages service is perfectly capable of managing other kinds
of information,
e.g.,
keeping track of machine-related infrastructural information;
however,
this is beyond the scope of the pilot.}

Naturally,
this raises questions as to the underlying technology which provides the
white pages service.
